---
title: "p8105_hw2_zv2138"
output: github_document
---

```{r Set Up}
library(tidyverse)
library(readxl)
```

# Problem 1

## Read and clean the Mr. Trash Wheel Sheet
```{r}
trash_wheel = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1, range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  filter(!(str_detect(month, "Total"))) %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0),
    month = str_to_lower(month),
    month = str_replace_all(month,"decemeber", "december")
  ) 
```

## Read and Clean Precipitation data for 2018 and 2019
```{r}
rain_2019 = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    mutate(
      year = 2019
    )
rain_2018 = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, range = "A2:B14",) %>% 
    janitor::clean_names() %>% 
    mutate(
       year = 2018
    )
```

## Combining Precipitation data sets 
```{r}
rain_data = bind_rows(rain_2018, rain_2019) %>% 
  mutate(
    month = str_to_lower(month.name[month])
  )
```

Mr. Trash Wheel filled `r nrow(trash_wheel)` dumpsters from `r min(pull(trash_wheel,date))` to `r max(pull(trash_wheel,date))`. In total `r sum(pull(trash_wheel, weight_tons))` tons of trash was collected during this time period. Mr. Trash Wheel collected various types of trash including `r colnames(trash_wheel[7:13])`. In fact the median number of sports balls that could be found in a dumpster in 2019 was `r median(pull(filter(trash_wheel,year == 2019), sports_balls))`.

Precipitation data was also collected over the years 2018 and 2019. Over the two years a total of `r sum(pull(rain_data, total))` inches of rain fell. In 2018 alone `r sum(pull(filter(rain_data, year == 2018), total))` inches of rain fell.

# Problem 2

## Read and Clean pols-month.csv
```{r}
pols_month = read_csv("./pols-month.csv", show_col_types = F) %>% 
  janitor::clean_names() %>% 
  mutate(mon = as.character(mon)) %>% 
  separate(mon, into = c("year","month", "day"), sep = "-") %>% 
  mutate(year = as.integer(year)) %>% 
  mutate(month = as.integer(month)) %>% 
  mutate(month = str_to_lower(month.name[month])) %>% 
  mutate(prez_dem = ifelse(prez_dem == 1, "dem", "")) %>% 
  mutate(prez_gop = ifelse(prez_gop == 1, "gop", "")) %>% 
  mutate(president = paste(prez_gop,prez_dem)) %>% 
  mutate(president = str_trim(president)) %>% 
  subset(select = -c(day, prez_gop, prez_dem))
```

## Read and Clean snp.csv
```{r}
snp = read_csv("./snp.csv", show_col_types = F) %>% 
  janitor::clean_names() %>% 
  mutate(date = as.character(date)) %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(year = as.integer(year)) %>% 
  mutate(year = (ifelse(year <= 15, year + 2000, year + 1900))) %>% 
  mutate(month = as.integer(month)) %>%
  mutate(month = str_to_lower(month.name[month])) %>% 
  subset(select = -(day)) %>% 
  relocate(year, month)
```

## Read and Clean unemployment.csv 
```{r}
unemployment = read_csv("./unemployment.csv", show_col_types = F) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "percent_unemployed"
  ) %>% 
  mutate(month = match(month, tolower(month.abb))) %>% 
  mutate(month = str_to_lower(month.name[month]))
```

## Merge unemployment, snp, and pols_month
```{r}
joined_df = left_join(pols_month, snp, by = c("year", "month"))
joined_df = left_join(joined_df, unemployment, by = c("year", "month"))   
```

Pols_month data file contains data on the number of national politicians (governors, senators, and presidents) who are democratic or republican at any given time. Pols_month has data on `r nrow(pols_month)` politicians. Data was collected from `r min(pull(pols_month,year))` to `r max(pull(pols_month,year))`. 

Snp data file contains data on the closing values of the Standard and Poor's stock market index for a given date. Snp has data from `r nrow(snp)` index observations. Data was collected from `r min(pull(snp,year))` to `r max(pull(snp,year))`. 

Unemployment data file contains data on the the percentage of unemployment on given year and month. The file contains `r nrow(unemployment)` data entries. Data was collected from `r min(pull(unemployment,year))` to `r max(pull(unemployment,year))`. 

The three data files were merged together to create the data set joined_df. This file contains `r nrow(joined_df)` observations. Data was collected from  `r min(pull(joined_df,year))` to `r max(pull(joined_df,year))`. 

# Problem 3

## Load and tidy the data - Popular Baby Names
```{r}
baby_names = read_csv("./Popular_Baby_Names.csv", show_col_types = F) %>% 
  janitor::clean_names() %>% 
  unique() %>% 
  mutate(ethnicity = str_replace_all(ethnicity,"ASIAN AND PACI$", "ASIAN AND PACIFIC ISLANDER")) %>% 
  mutate(ethnicity = str_replace_all(ethnicity,"BLACK NON HISP$", "BLACK NON HISPANIC")) %>% 
  mutate(ethnicity = str_replace_all(ethnicity,"WHITE NON HISP$", "WHITE NON HISPANIC"))


olivia = filter(baby_names, childs_first_name == "Olivia", gender == "FEMALE")
pivot_longer(olivia, !ethnicity, names_to = "year_of_birth", values_to = "rank")

```

