---
title: "p8105_hw2_zv2138"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

## Read and clean the Mr. Trash Wheel Sheet
```{r}
trash_wheel = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0)
  ) %>% 
  filter(!(str_detect(month, "Total")))
```

## Read and Clean Precipitation data for 2018 and 2019
```{r}
rain_2019 = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 4, range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    na.omit %>% 
    mutate(
      year = 2019
    )
rain_2018 = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, range = "A2:B14",) %>% 
    janitor::clean_names() %>% 
    mutate(
       year = 2018
    )
```

## Combining Precipitation data sets 
```{r}
rain_data = bind_rows(rain_2018, rain_2019) %>% 
  mutate(
    month = month.name[month]
  )
```

Mr. Trash Wheel filled `r nrow(trash_wheel)` dumpsters from `r min(pull(trash_wheel,date))` to `r max(pull(trash_wheel,date))`. In total `r sum(pull(trash_wheel, weight_tons))` tons of trash was collected during this time period. Mr. Trash Wheel collected various types of trash including `r colnames(trash_wheel[7:13])`. In fact the median number of sports balls that could be found in a dumpster in 2019 was `r median(pull(filter(trash_wheel,year == 2019), sports_balls))`.

Precipitation data was also collected over the years 2018 and 2019. Over the two years a total of `r sum(pull(rain_data, total))` inches of rain fell. In 2018 alone `r sum(pull(filter(rain_data, year == 2018), total))`.


# Problem 2